Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-19 writeup/gz_hubble_data.tex
splitting figure 9 into 2 figures

\begin{figure*}
\centering
\subfigure{[a]\label{fig:bias_a}\includegraphics[width=0.5\textwidth]{figures/p_vs_p_SB_redshift.pdf}}
\subfigure{[b]\label{fig:bias_b}\includegraphics[width=0.4\textwidth]{figures/z1_mu20_subplot1.pdf}}
\\
\subfigure{[c]\label{fig:bias_c}\includegraphics[width=0.5\textwidth]{figures/orangebars.pdf}}
\subfigure{[d]\label{fig:bias_d}\includegraphics[width=0.4\textwidth]{figures/z1_mu20_subplot2.pdf}}
\caption{Effects of redshift bias in 3,950 images in the \ferengi{} sample.
[a]: Each point in a given redshift and surface brightness bin represents a
unique galaxy. On the $y$-axis in each bin is the \ffeatures{} value of the
image of that galaxy redshifted to the value corresponding to that redshift
bin. On the $x$-axis is the \ffeatures{} value of the image of the same galaxy
redshifted to $z=0.3$. The dashed black lines represent the best-fit
polynomials to the data in each square. The solid black line represents
\ffeaturesz=\ffeaturesrest. Regions in which there is a single-valued
relationship between \ffeatures{} at high redshift and at $z=0.3$ are white;
those in which there is not are blue, and those with not enough data ($N<5$)
are gray. [b]: A larger version of the dark-outlined square in [a], containing
\ferengi{} galaxies that have been artificially redshifted to $z=1.0$ and have
surface brightnesses between $20.3 < \mu < 21.0$ $\rm (mag/arcsec^2)$. [c]: The
same data as [a] is shown. Each $z,\mu$ bin is divided into 4 sub-bins to
determine the range of intrinsic \ffeaturesrest{} for a given range of observed
\ffeaturesz{} values. In each sub-bin, the orange bars represent the inner
80$^\mathrm{th}$ percentiles of the data, the boundaries of which are the lower
and upper limits of the debiased values. [d]: The same data as [b], but
highlighting the upper and lower limit regions.}
\label{fig:f_vs_f}
\end{figure*}

@@

Figure~\ref{fig:f_vs_f} shows that the relationship between \ffeaturesz{} and
\ffeaturesrest{} is \emph{not} always single valued; hence, it is not
appropriate to correct galaxies that lie in certain regions of surface
brightness/redshift/\ffeatures{} space. These regions tend to have low
\ffeatures{} values at high redshift, but a wide range of values at $z=0.3$.
These regions contain two morphological types of galaxies: the first set are
genuine ellipticals, which have low values of \ffeatures{} at both high and low
redshift. The second group are disks whose features become indistinct at high
redshift; hence their \ffeatures{} value at $z=0.3$ may be quite high, while
the value observed at high redshift is very low. This effect is strongest at
high $z$ and low $\mu$, where features become nearly impossible to discern in
the images (see top right panel of Figure~\ref{fig:bias_a}).

The criteria for determining whether a region of this space is single-valued,
and therefore correctable, is as follows: In each surface brightness and
redshift bin, the relationship between \ffeaturesz{} and \ffeaturesrest{} is
modeled by fitting the data with a polynomial of degrees $n=3$, 2, and 1, and
using the best formal fit out of the three as measured by the sum of the
residuals. These fits are shown as the dashed black lines in
Figure~\ref{fig:bias_a}. Any flat regions of the polynomial fits are areas in which
there is not a clear single-valued relationship between \ffeaturesz{} and
\ffeaturesrest; this is quantified by setting a minimum slope cut of 0.4. Any
data in which the polynomial fit has a slope less than this value is considered
\emph{not} one-to-one, and therefore \ffeaturesz{} cannot be boosted to its
\ffeaturesrest{} value.  Galaxies in this region are referred to as the
\emph{lower limit} sample, because the most stringent correction available is
that the weighted \ffeatures{} is a lower limit to the true value.  These
regions are highlighted in blue in Figure~\ref{fig:bias_a}. Uncolored (white)
regions of the plot have sufficiently high slopes to consider the relationship
as single-valued; galaxies in these regions are considered ``correctable'', and
only these are used in measuring the parameters for the $\zeta$ function
(Section~\ref{ssec:zeta}). Only surface brightness/redshift bins with at least
5 galaxies were considered; regions with fewer than 5 galaxies are considered
to have ``not enough information'' to determine the \ffeaturesz{} and
\ffeaturesrest{} relationship, colored gray in Figure~\ref{fig:bias_a}.

The unshaded regions in Figure~\ref{fig:bias_a} define discrete ranges of redshift,
surface brightness, and \ffeatures{} a galaxy must have in order for the
debiased correction to be confidently applied to a galaxy in the GZH sample.
While the appropriate correctable regions were defined discretely, the true
correctable region is assumed to be a smooth function of $z$, $\mu$, and
\ffeatures{}. To define this smooth space, we calculate the shape of the convex
hull that encloses the correctable and lower-limit \ferengi{} galaxies in
$z$-$\mu$-\ffeatures{} space. The boundaries are then adjusted until the
contamination from both groups is minimized. The resulting hulls define the
correctable and lower-limit regions for categorizing the \hst{} galaxies. The
results of this method and final categorization of the \hst{} sample are in
Table~\ref{tbl:hubble_debiasable}. Of the galaxies at redshift higher than
$z=0.3$, 17\% can be debiased using the $\zeta$ method, 27\% cannot be
debiased, and 56\% cannot have the potential for a correction determined, due
either to an unknown redshift or insufficient \ferengi{} images corresponding
to the relevant $z$-$\mu$ values.

For the ``lower-limit'' galaxies for which a single debiased \ffeatures{} value
cannot be confidently assigned, the \emph{range} of debiased values is
estimated using a method visualized in Figure~\ref{fig:bias_c}. This uses the
\ferengi{} simulated data to analyze the range of intrinsic \ffeaturesrest{}
values for any given observed \ffeatures{} value, again as a function of
surface brightness and redshift. Each $z$,$\mu$ bin, shows the spread of
intrinsic values of \ffeaturesrest{} for four ranges of observed \ffeatures.
The range of intrinsic values is defined as the inner 80\% of the data,
represented by the orange bars in Figure~\ref{fig:bias_c}. For any galaxy which
cannot be directly debiased by the $\zeta$ method, these ranges are used to
denote the upper and lower limits on the expected values \ffeaturesrest{} as a
function of the observed \ffeatures.

@@

\begin{figure*}
\centering
\subfigure{[a]\label{a}\includegraphics[width=0.5\textwidth]{figures/p_vs_p_SB_redshift_bar.pdf}}
\subfigure{[b]\label{b}\includegraphics[width=0.4\textwidth]{figures/z1_mu20_subplot1_bar.pdf}}
\\
\subfigure{[c]\label{c}\includegraphics[width=0.5\textwidth]{figures/orangebars_bar.pdf}}
\subfigure{[d]\label{d}\includegraphics[width=0.4\textwidth]{figures/z1_mu20_subplot2_bar.pdf}}
\caption{Similar to Figure~\ref{fig:f_vs_f}, showing effects of redshift bias
for 3,950~images in the \ferengi{} sample. [a]: Each point in a given redshift
and surface brightness bin represents a unique galaxy. The $y$-axis in each
bin is the \ffeatures{} value of the image of that galaxy redshifted to the
value corresponding to that redshift bin. The $x$-axis is the \ffeatures{}
value of the image of the same galaxy redshifted to $z=0.3$. The dashed
black lines represent the best-fit polynomials to the data in each square.
The solid black line represents \ffeaturesz=\ffeaturesrest. Regions in
which there is a single-valued relationship between \ffeatures{} at high
redshift and at $z=0.3$ are marked in white; those in which there is not
are blue, and those with not enough data ($N<5$) are gray. The mean
normalized $\chi^2$ of the change in vote fraction is 0.08. None of the
correlations for higher-order tasks (including bars) were applied to
galaxies in the GZH catalog. [b]: A larger version of the dark-outlined
square in [a], containing \ferengi{} galaxies artificially redshifted to
$z=1.0$ and have surface brightnesses between $20.3 < \mu < 21.0$ $\rm
(mag/arcsec^2)$. [c]: The same data as [a]. Each $z,\mu$ bin is divided
into 4~sub-bins to determine the range of intrinsic \ffeaturesrest{} for a
given range of observed \ffeaturesz{} values. In each sub-bin, the orange
bars represent the inner 80$^\mathrm{th}$ percentiles of the data, the
boundaries of which are the lower and upper limits of the debiased values.
[d]: The same data as [b], but highlighting the upper and lower limit
regions.}
\label{fig:f_vs_fbar}
\end{figure*}

>>>>

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/p_vs_p_sb_redshift.pdf}
\caption{Effects of redshift bias in 3,950 images in the \ferengi{} sample.
Each point in a given redshift and surface brightness bin represents a
unique galaxy. On the $y$-axis in each bin is the \ffeatures{} value of the
image of that galaxy redshifted to the value corresponding to that redshift
bin. On the $x$-axis is the \ffeatures{} value of the image of the same galaxy
redshifted to $z=0.3$. The dashed black lines represent the best-fit
polynomials to the data in each square. The solid black line represents
\ffeaturesz=\ffeaturesrest. Regions in which there is a single-valued
relationship between \ffeatures{} at high redshift and at $z=0.3$ are white;
those in which there is not are blue, and those with not enough data ($N<5$)
are gray. A larger version of the bin outlined at $z=1.0$ and $20.3 < \mu < 21.0$ $\rm (mag/arcsec^2)$ is shown in Figure~\ref{fig:f_vs_f_zoom}.}
\label{fig:f_vs_f}
\end{figure*}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/z1_mu20_subplot2.pdf}
\caption{A larger version of the dark-outlined square in Figure~\ref{fig:f_vs_f}, containing
\ferengi{} galaxies that have been artificially redshifted to $z=1.0$ and have
surface brightnesses between $20.3 < \mu < 21.0$ $\rm (mag/arcsec^2)$. The orange bars represent the inner
80$^\mathrm{th}$ percentiles of the data, the boundaries of which are the lower
and upper limits of the debiased values.}
\label{fig:f_vs_f_zoom}
\end{figure}

@@

Figure~\ref{fig:f_vs_f} shows that the relationship between \ffeaturesz{} and
\ffeaturesrest{} is \emph{not} always single valued; hence, it is not
appropriate to correct galaxies that lie in certain regions of surface
brightness/redshift/\ffeatures{} space. These regions tend to have low
\ffeatures{} values at high redshift, but a wide range of values at $z=0.3$.
These regions contain two morphological types of galaxies: the first set are
genuine ellipticals, which have low values of \ffeatures{} at both high and low
redshift. The second group are disks whose features become indistinct at high
redshift; hence their \ffeatures{} value at $z=0.3$ may be quite high, while
the value observed at high redshift is very low. This effect is strongest at
high $z$ and low $\mu$, where features become nearly impossible to discern in
the images (see top right panel of Figure~\ref{fig:f_vs_f}).

The criteria for determining whether a region of this space is single-valued,
and therefore correctable, is as follows: In each surface brightness and
redshift bin, the relationship between \ffeaturesz{} and \ffeaturesrest{} is
modeled by fitting the data with a polynomial of degrees $n=3$, 2, and 1, and
using the best formal fit out of the three as measured by the sum of the
residuals. These fits are shown as the dashed black lines in
Figure~\ref{fig:f_vs_f}. Any flat regions of the polynomial fits are areas in which
there is not a clear single-valued relationship between \ffeaturesz{} and
\ffeaturesrest; this is quantified by setting a minimum slope cut of 0.4. Any
data in which the polynomial fit has a slope less than this value is considered
\emph{not} one-to-one, and therefore \ffeaturesz{} cannot be boosted to its
\ffeaturesrest{} value.  Galaxies in this region are referred to as the
\emph{lower limit} sample, because the most stringent correction available is
that the weighted \ffeatures{} is a lower limit to the true value.  These
regions are highlighted in blue in Figure~\ref{fig:f_vs_f}. Uncolored (white)
regions of the plot have sufficiently high slopes to consider the relationship
as single-valued; galaxies in these regions are considered ``correctable'', and
only these are used in measuring the parameters for the $\zeta$ function
(Section~\ref{ssec:zeta}). Only surface brightness/redshift bins with at least
5 galaxies were considered; regions with fewer than 5 galaxies are considered
to have ``not enough information'' to determine the \ffeaturesz{} and
\ffeaturesrest{} relationship, colored gray in Figure~\ref{fig:f_vs_f}.

The unshaded regions in Figure~\ref{fig:f_vs_f} define discrete ranges of redshift,
surface brightness, and \ffeatures{} a galaxy must have in order for the
debiased correction to be confidently applied to a galaxy in the GZH sample.
While the appropriate correctable regions were defined discretely, the true
correctable region is assumed to be a smooth function of $z$, $\mu$, and
\ffeatures{}. To define this smooth space, we calculate the shape of the convex
hull that encloses the correctable and lower-limit \ferengi{} galaxies in
$z$-$\mu$-\ffeatures{} space. The boundaries are then adjusted until the
contamination from both groups is minimized. The resulting hulls define the
correctable and lower-limit regions for categorizing the \hst{} galaxies. The
results of this method and final categorization of the \hst{} sample are in
Table~\ref{tbl:hubble_debiasable}. Of the galaxies at redshift higher than
$z=0.3$, 17\% can be debiased using the $\zeta$ method, 27\% cannot be
debiased, and 56\% cannot have the potential for a correction determined, due
either to an unknown redshift or insufficient \ferengi{} images corresponding
to the relevant $z$-$\mu$ values.

For the ``lower-limit'' galaxies for which a single debiased \ffeatures{} value
cannot be confidently assigned, the \emph{range} of debiased values is
estimated using a method visualized in Figure~\ref{fig:f_vs_f_zoom}. This uses the
\ferengi{} simulated data to analyze the range of intrinsic \ffeaturesrest{}
values for any given observed \ffeatures{} value, again as a function of
surface brightness and redshift. In each unclassifiable (shaded blue) $z$,$\mu$ bin, the spread of
intrinsic values of \ffeaturesrest{} for four ranges of observed \ffeatures is computed.
The range of intrinsic values is defined as the inner 80\% of the data,
represented by the orange bars in Figure~\ref{fig:f_vs_f_zoom}. For any galaxy which
cannot be directly debiased by the $\zeta$ method, these ranges are used to
denote the upper and lower limits on the expected values \ffeaturesrest{} as a
function of the observed \ffeatures.

@@

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{figures/p_vs_p_SB_redshift_bar.pdf}
\caption{Similar to Figure~\ref{fig:f_vs_f}, showing effects of redshift bias
for 3,950~images in the \ferengi{} sample. Each point in a given redshift
and surface brightness bin represents a unique galaxy. The $y$-axis in each
bin is the \ffeatures{} value of the image of that galaxy redshifted to the
value corresponding to that redshift bin. The $x$-axis is the \ffeatures{}
value of the image of the same galaxy redshifted to $z=0.3$. The dashed
black lines represent the best-fit polynomials to the data in each square.
The solid black line represents \ffeaturesz=\ffeaturesrest. Regions in
which there is a single-valued relationship between \ffeatures{} at high
redshift and at $z=0.3$ are marked in white; those in which there is not
are blue, and those with not enough data ($N<5$) are gray. A larger version of the bin outlined at $z=1.0$ and $20.3 < \mu < 21.0$ $\rm (mag/arcsec^2)$ is shown in Figure~\ref{fig:f_vs_fbar_zoom}. The mean
normalized $\chi^2$ of the change in vote fraction is 0.08. None of the
correlations for higher-order tasks (including bars) were applied to
galaxies in the GZH catalog.}
\label{fig:f_vs_fbar}
\end{figure*}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/z1_mu20_subplot2_bar.pdf}
\caption{A larger version of the dark-outlined square in Figure~\ref{fig:f_vs_fbar}, containing
\ferengi{} galaxies that have been artificially redshifted to $z=1.0$ and have
surface brightnesses between $20.3 < \mu < 21.0$ $\rm (mag/arcsec^2)$. The orange bars represent the inner
80$^\mathrm{th}$ percentiles of the data, the boundaries of which are the lower
and upper limits of the debiased values.}
\label{fig:f_vs_fbar_zoom}
\end{figure}
